\section{Introduction}
\label{sec:intro}

Quantum computing offers a computational model distinct from classical computing by operating on qubits rather than binary bits. Each qubit is represented as a quantum state with probability amplitudes (e.g., $|\psi\rangle = \alpha|0\rangle + \beta|1\rangle$), where superposition enables multiple states, and entanglement creates correlations between qubits. These properties enable parallelism beyond classical computation~\cite{miguel2023enhancing, renner2022computational}. Despite these advantages, current quantum computers remain limited to Noisy Intermediate-Scale Quantum (NISQ) devices with high error rates and short coherence times~\cite{suzuki2022quantum, preskill2019quantum}. To overcome these limitations, classical hardware is used for full state vector simulation, which evaluates all amplitudes and provides exact results for algorithm validation~\cite{fu2024achieving, pan2022simulation}.

While classical simulation guarantees exact results, transitioning quantum applications from research to practical environments introduces a deployment bottleneck. Emerging workloads, such as iterative variational quantum algorithms (e.g., VQE) and on-device quantum machine learning, demand real-time responsiveness and bounded per-iteration latency. To satisfy these constraints, systems must execute simulations locally on resource-constrained edge devices rather than offloading them to remote cloud datacenters, which incurs network round-trip overhead on every iteration. However, enabling local execution is obstructed by an exponential memory wall: the state vector footprint doubles with each additional qubit. While server-grade discrete GPUs (e.g., NVIDIA A100) provide dedicated VRAM (80GB) to support this scaling, standard edge systems (e.g., NVIDIA Jetson Orin Nano) impose a physical limit of only 8GB of shared DRAM. Because evaluating intermediate-scale algorithms demands hundreds of gigabytes to terabytes of memory, this disparity between massive data and limited edge hardware makes local simulation infeasible.



To overcome this memory barrier, existing approaches attempt to deploy high-performance frameworks (e.g., cuQuantum~\cite{bayraktar2023cuquantum}) using out-of-core memory management. However, applying these frameworks to edge environments exposes an architecture mismatch. In contrast to discrete GPUs, edge devices employ a Unified Memory Architecture (UMA) where the CPU and GPU share a constrained physical memory pool. When standard frameworks attempt to allocate large state vectors within this shared pool, they exhaust the available DRAM. Lacking edge-aware memory management, they suffer from severe execution bottlenecks as memory pressure builds, exhausting available GPU memory and triggering allocation failures. In contrast, \EdgeQuantum leverages a UVM-aware tiered memory architecture to bypass these hardware bottlenecks, sustaining stable execution as the state vector exceeds physical memory limits.
%To overcome this memory barrier, existing approaches attempt to deploy high-performance frameworks (e.g., cuQuantum~\cite{bayraktar2023cuquantum}) using out-of-core memory management. However, applying these frameworks to edge environments exposes an architecture mismatch. In contrast to discrete GPUs, edge devices employ a Unified Memory Architecture (UMA) where the CPU and GPU share a constrained physical memory pool. When standard frameworks attempt to allocate large state vectors within this shared pool, they exhaust the available DRAM. Lacking edge-aware memory management, they fall back on operating system virtual memory (OS-native swapping) to spill data to external NVMe storage, which triggers I/O thrashing. Figure~\ref{intro_motivation_b} shows the impact of this mismatch on a Variational Quantum Eigensolver (VQE) workload. CPU-only execution (Cirq) avoids GPU memory exhaustion but is computationally expensive, requiring over 100 seconds (109,168ms) for a single 29-qubit iteration. Conversely, GPU-accelerated execution (cuQuantum) degrades as it triggers OS swapping near physical memory limits (96.6ms at 29 qubits), before crashing with an Out-of-Memory (OOM) error at 30 qubits due to GPU workspace overheads and CUDA allocator constraints. In contrast, \EdgeQuantum leverages a UMA-optimized architecture to bypass OS bottlenecks, maintaining millisecond-level latency (2.0~ms) at 29 qubits and completing the 30-qubit out-of-core simulation in 10.3 seconds.





\begin{table}[t]
\caption{Comparison with prior work across key capabilities (Tiered: Tiered Memory Support, Comp: Compression Support).}
\centering
\scriptsize
\begin{tabularx}{\columnwidth}{p{2cm}|c|c|c|c|c}
\toprule
\textbf{Framework} & \textbf{Target} & \textbf{GPU} & \textbf{Tiered} & \textbf{Comp} & \textbf{Max Qubit} \\
\midrule
Qiskit Aer~\cite{qiskit} & General & \checkmark &  &  & 29 \\
cuQuantum~\cite{bayraktar2023cuquantum} & General & \checkmark & \checkmark &  & 40+ \\
PennyLane~\cite{bergholm2018pennylane} & General & \checkmark &  &  & 25+ \\
HyQuas~\cite{zhang2021hyquas} & HPC & \checkmark &  &  & 36 \\
Atlas~\cite{xu2024atlas} & HPC & \checkmark & \checkmark  &  & 36 \\
ScaleQsim~\cite{10.1145/3771577} & HPC & \checkmark &  &  & 42 \\
BMQSim~\cite{zhang2025bmqsim} & Datacenter & \checkmark & \checkmark & \checkmark & 36 \\
\textbf{\EdgeQuantum} & \textbf{Edge/IoT} & \checkmark & \checkmark & \checkmark & \textbf{39} \\
\bottomrule
\end{tabularx}
\label{intro_table}
\end{table}

Many previous studies, as summarized in Table~\ref{intro_table}, optimize quantum circuit simulation for state-of-the-art computing infrastructure, including datacenter and HPC systems. Frameworks such as Qiskit Aer~\cite{qiskit} and PennyLane~\cite{bergholm2018pennylane} target workstations and assume the state vector fits in physical GPU VRAM and DRAM. Several works, such as HyQuas~\cite{zhang2021hyquas}, Atlas~\cite{xu2024atlas}, ScaleQsim~\cite{10.1145/3771577}, and SnuQS~\cite{park2022snuqs} use multi-node execution to scale memory capacity but rely on high-bandwidth interconnects such as NVLink and InfiniBand and large GPU clusters. Other works, such as cuQuantum~\cite{bayraktar2023cuquantum} and BMQSim~\cite{zhang2025bmqsim} use tiered memory or data compression but target server-grade GPUs with dedicated VRAM. These approaches do not consider the hardware characteristics of edge systems, including limited NVMe bandwidth and shared memory constraints in Unified Virtual Memory (UVM) architectures. Deploying these frameworks on edge devices leads to Out-of-Memory failures or performance degradation due to I/O contention.

In this paper, we present \EdgeQuantum, a scalable full state-vector simulation framework for resource-constrained edge devices. While existing out-of-core simulators address memory limitations through OS-native swapping or server-grade abstractions, they do not coordinate the end-to-end I/O path for UMA environments. \EdgeQuantum integrates the unified physical DRAM pool and commodity NVMe storage into a 2-tier memory architecture that decouples logical state-vector management from physical residency. Specifically, \EdgeQuantum introduces 1) state vector partitioning across shared DRAM and NVMe to enable large-scale simulations beyond hardware limits, 2) an edge-aware asynchronous Unified Virtual Memory (UVM) pipeline that serializes storage access to avoid I/O contention and overlaps flash operations with GPU computation, and 3) an on-the-fly LZ4 ping-pong compression strategy that expands effective storage capacity by 128$\times$ (with a 256$\times$ per-file LZ4 ratio) while masking NVMe latency. Our evaluation on an 8~GB Jetson Orin Nano demonstrates that \EdgeQuantum successfully performs a 39-qubit simulation (4~TB state vector footprint), achieving a 512$\times$ capacity expansion over the physical memory limit. We open-source \EdgeQuantum at \url{https://github.com/sunggonkim/IoTJ-EdgeQuantum}.



%Specifically, \EdgeQuantum introduces 1) state vector partitioning across shared DRAM and NVMe to enable large-scale simulations beyond hardware limits, 2) an edge-aware asynchronous Unified Virtual Memory (UVM) pipeline that serializes storage access to avoid I/O contention and overlaps flash operations with GPU computation, and 3) an on-the-fly LZ4 ping-pong compression strategy that bypasses the impossibility of in-place updates for variable-sized data by alternating backing files, thereby expanding effective storage capacity by $242.7\times$ while masking NVMe latency.