\section{Related Work}
\label{sec:related}
\noindent\textbf{Optimizing Quantum Circuit Simulation.} Previous studies on quantum circuit simulation have focused on utilizing massive parallelism in GPUs and distributed clusters for full state vector simulation. These approaches accelerate execution by distributing the state across aggregated VRAM in leadership-class supercomputers~\cite{cuquantum, qsim, guerreschi2020intel, villalonga2019flexible}. Other studies propose static compilation and circuit partitioning to reduce communication overhead. These methods maximize kernel occupancy through precompiled kernels and optimized execution plans~\cite{zhang2021hyquas, xu2024atlas, zhang2022uniq}. Additionally, tensor network-based approaches reduce memory footprint by decomposing quantum states into interconnected tensors. These approaches provide savings for low-entanglement circuits while facing exponential complexity growth in deep, highly entangled circuits~\cite{gray2018quimb, lykov2022tensor, pan2022simulation, biamonte2017tensor}.

Our study aligns with prior efforts to improve simulation efficiency. However, existing datacenter-centric approaches such as \textit{ScaleQsim} and \textit{cuQuantum} assume abundant memory resources and high-bandwidth interconnects. This assumption leads to execution failures on embedded edge devices where physical memory is strictly limited (e.g., 8~GB). In contrast, \EdgeQuantum targets resource-constrained devices. It partitions the state vector across the storage hierarchy rather than network nodes. This design uses a specialized 4-buffer ring pipeline.

\noindent\textbf{Memory Expansion and I/O Optimization Strategies.} Previous studies on memory expansion focus on extending simulation capacity by offloading state vectors to high-performance NVMe SSDs in server environments. These approaches rely on OS-level demand paging or explicit memory mapping~\cite{park2022snuqs, wang2023enabling, reano2017intra}. Other work proposes lossless compression and adaptive encoding to reduce memory consumption. These methods trade CPU cycles for capacity~\cite{zhang2025bmqsim, wu2019full, zhang2024overcoming, li2021sv}. In the broader context of deep learning, approaches such as vDNN and FlashNeuron simplify memory management by offloading intermediate feature maps to backing stores~\cite{rhu2016vdnn, bae2021flashneuron}.

While our study aligns with these offloading and compression techniques, generic strategies remain inefficient on edge devices due to the bandwidth asymmetry of consumer-grade SSDs and UMA constraints. Prior server-centric frameworks~\cite{zhang2025bmqsim, wu2019full, li2021sv} primarily target high-end infrastructure and do not address the variable-size update constraints of commodity edge storage. \EdgeQuantum employs an asynchronous I/O strategy that serializes read and write operations to avoid NVMe contention. This is a critical optimization rarely considered in server-centric storage engines. Combined with UVM-aware zero-copy buffering and a dual-file ping-pong compression strategy, \EdgeQuantum improves pipeline efficiency and minimizes CPU-GPU synchronization overhead.

\begin{comment}
  
\subsection{Optimizing Quantum Circuit Simulation}


There have been many studies that optimize quantum circuit simulation to enhance performance on high-end hardware.
Previous studies~\cite{cuquantum, qsim, guerreschi2020intel, villalonga2019flexible} focused on utilizing massive parallelism in GPUs and distributed clusters for full state vector simulation.
These approaches accelerate execution and extend scalability by distributing the state vector across aggregated video memory (VRAM) in leadership-class supercomputers.
Other studies~\cite{zhang2021hyquas, xu2024atlas, zhang2022uniq} have proposed static compilation and circuit partitioning techniques to reduce communication overhead.
These methods analyze quantum circuits in advance to generate optimized execution plans or precompiled kernels, aiming to maximize kernel occupancy and minimize inter-node data transfer.
In addition, tensor network-based approaches have been proposed~\cite{gray2018quimb, lykov2022tensor, pan2022simulation, biamonte2017tensor} to reduce the memory footprint by decomposing quantum states into interconnected tensors.
These methods provide significant memory savings for shallow or low-entanglement circuits but face exponential complexity growth when simulating deep circuits with high entanglement.

Our study aligns with these prior efforts in improving the computational efficiency of quantum circuit simulation.
However, \EdgeQuantum aims to democratize quantum simulation by targeting resource-constrained edge devices rather than relying on inaccessible HPC infrastructure.
Existing datacenter-centric approaches such as \textit{ScaleQsim} and \textit{cuQuantum} assume abundant memory resources and high-bandwidth interconnects, which leads to execution failures on embedded devices where physical memory is strictly limited (e.g., 8 GB).
Through a tiered-memory architecture, \EdgeQuantum partitions the state vector across the storage hierarchy rather than across network nodes, and employs a specialized 4-buffer pipeline.
This allows \EdgeQuantum to enable large-scale simulation on commodity hardware without requiring millions of dollars in infrastructure.

\subsection{Memory Expansion and I/O Optimization Strategies}

To address the physical memory bottleneck, several simulation frameworks and system-level optimizations have been proposed to utilize secondary storage or compression.
Previous studies~\cite{park2022snuqs, wang2023enabling, reano2017intra} focused on extending simulation capacity by offloading state vectors to high-performance NVMe SSDs in server environments.
These approaches manage the movement of data pages between host memory and storage, often relying on OS-level demand paging or explicit memory mapping (mmap) to support simulations exceeding physical RAM capacity.
Other studies~\cite{zhang2025bmqsim, wu2019full, zhang2024overcoming, li2021sv} have proposed lossless compression and adaptive error-bounded encoding techniques to reduce memory consumption.
These methods dynamically compress state vectors during simulation, allowing systems to store larger quantum states within limited memory footprints by trading CPU cycles for capacity.
In the broader context of deep learning, approaches such as vDNN~\cite{rhu2016vdnn} and FlashNeuron~\cite{bae2021flashneuron} simplify memory management by offloading intermediate feature maps to backing stores.

These approaches highlight key techniques for improving capacity in data-intensive applications.
Similarly, \EdgeQuantum faces comparable challenges in edge quantum simulation, where the state vector size far exceeds the unified memory capacity of the SoC.
However, applying generic offloading techniques to edge devices is inefficient due to the bandwidth asymmetry of consumer-grade SSDs and the specific constraints of the Unified Memory Architecture (UMA).
To address this, \EdgeQuantum employs an asynchronous I/O strategy that serializes read and write operations to avoid NVMe contention, a critical optimization ignored by server-centric storage engines.
Combined with UVM-aware zero-copy buffering and adaptive LZ4 compression, \EdgeQuantum improves pipeline efficiency while minimizing CPU-GPU synchronization overhead.  
\end{comment}